### 2021-02-01 research project/sections/Analysis_of_results.tex
research project

\begin{tikzpicture}[scale=1.1]
    \begin{axis}[
        title={Comparing the time taken for approximating the value of $\pi$},
        xlabel={Decimal places},
        ylabel={Time taken [ms]},
        xmin=0, xmax=60,
        ymin=0, ymax=10,
        ytick={0,2,4,6,8,10},
        xtick={0,20,40,60},
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        scale=1.2
    ]

@@

The results gathered from this experiment are descisive.

>>>>

\begin{tikzpicture}[scale=1.1]
    \begin{axis}[
        title={Comparing the time taken for approximating the value of $\pi$},
        xlabel={Decimal places},
        ylabel={Time taken [ms]},
        xmin=0, xmax=60,
        ymin=0, ymax=10,
        ytick={0,2,4,6,8,10},
        xtick={0,10,20,30,40,50,60},
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        scale=1.2
    ]

@@

The results gathered from this experiment show linear relationship between the amount of decimal 
places approximated and the time required for this approximation. The two lines of best fit for 
the two methods show similarities but also differences. Firstly, from 5 to 20 decimal places, 
Madhava's

### 2021-02-01 research project/sections/Background.tex
rp
research project

Madhava is the first mathematician known to have found the series notation for 
the inverse tangent function, in the Kerala school of mathematics of Medieval India 
\cite{amermathmonthly}. The research made at the Kerala school was documented by 
many mathematicians of their time, namely the astronomer-mathematician Jyesthadeva 
in his treatise \textit{Yukti-Bhasa} written in the Malayalam language \cite{oconnor_robertson_2000}. 
Madhava's infinite series expansion of the inverse tangent function can also be 
found in this treatise. R. C. Gupta, in his translation, describes this expansion as one 
where the arc $\theta$ is equal to the sum of a first term, "the product of the given sine 
and radius of the desired arc divided by the cosine of the arc" \footnotemark, followed by terms that "are 
obtained by a process of iteration": . \cite{rc_gupta_mgseries}

@@

Which has come to be known as the Leibniz formula for $\pi$, after the German mathematician 
who discovered the same formula two decades later. \cite{edwards_1994}

@@

François Viète approached the value of $\pi$ from a 
geometric standpoint, and found an infinite product. 
He was able to calculate $\pi$ to a place of 9 decimal points, in 
the year 1593 \cite{Kreminski}, using his method. His method is reminiscent 
of Archimedes' method, where the length of a side is calculated \cite{archimedespi}, 
but differs in that it consists of finding the area of a polygon of $n$ sides in a circle of constant 
radius, rather than the circumference. As the value of $n$ is increased, the area of the $n$-gon 
tends toward the area of a circle. The geometric origin of this formula 
can be found using simple right-angle trigonometry, by first finding 
the lengths $OH$ and subsequently $BD$ in (Figure 1 ) \ref{fig:image1}.
% $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
With the radius of the circle with center $R = OB$,

@@

>>>>

Madhava is the first mathematician known to have found the series notation for 
the inverse tangent function, in the Kerala school of mathematics of Medieval India 
\cite{amermathmonthly}. The research made at the Kerala school was documented by 
many mathematicians of their time, namely the astronomer-mathematician Jyesthadeva 
in his treatise \textit{Yukti-Bhasa} written in the Malayalam language \cite{oconnor_robertson_2000}. 
Madhava's infinite series expansion of the inverse tangent function can also be 
found in this treatise. R. C. Gupta, in his translation, describes this expansion as one 
where the arc $\theta$ is equal to the sum of a first term, "the product of the given sine 
and radius of the desired arc divided by the cosine of the arc"\footnotemark, followed by terms that "are 
obtained by a process of iteration": in which the original term is multiplied by the 
square of the sine and divided by the square of the cosine. Thereafter, each term is 
divided by an odd number in order (1, 3, 5, 7, \dots). It is then said that the arc can 
be found by "adding and subtracting respectively the terms of odd rank and those of even rank",  
definition of an alternating series. \cite{rc_gupta_mgseries}

@@

Which has come to be known as the Leibniz formula for $\pi$, after the German mathematician 
who discovered the same formula two decades later. \cite{edwards_1994} This series converges to $\pi$, as seen in (Figure 1) \ref{fig:numberline}

@@

François Viète approached the value of $\pi$ from a 
geometric standpoint, and found an infinite product. 
He was able to calculate $\pi$ to a place of 9 decimal points, in 
the year 1593 \cite{Kreminski}, using his method. His method is reminiscent 
of Archimedes' method, where the length of a side is calculated \cite{archimedespi}, 
but differs in that it consists of finding the area of a polygon of $n$ sides in a circle of constant 
radius, rather than the circumference. As the value of $n$ is increased, the area of the $n$-gon 
tends toward the area of a circle. The geometric origin of this formula 
can be found using simple right-angle trigonometry, by first finding 
the lengths $OH$ and subsequently $BD$ in (Figure 2) \ref{fig:image1}.
% $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
With the radius of the circle with center $R = OB$,

@@

\begin{figure}[h]

    \begin{tikzpicture}
        \draw[latex-latex] (-0.5,0) -- (4.5,0);% the x-axis
        \foreach \x in  {0,...,4} % tick marks
          \draw (\x,0) -- (\x,-3pt);
        \foreach \x in {0,...,4} % the numbers
          \node [below] at (\x,-0.1) {$\x$};
        \node [below] at (3.14,0.5) {$\pi$};
        \draw (3.14,0) -- (3.14,3pt);

        \draw[->, dotted] (3.14,0) -- (3.14,-3);% the x-axis

        \draw[xshift=-1cm, dotted] (1,-0.7) node[circle,fill,inner sep=1pt](a){} -- (5,-0.7) 
node[circle,fill,inner sep=1pt,label=right:$+\frac{4}{1}$](b){};

\draw[xshift=-1cm, dotted] (3.67,-1.1) node[circle,fill,inner sep=1pt, label=left:$-\frac{4}{3}$](c){} -- (5,-1.1) 
node[circle,fill,inner sep=1pt](d){};

\draw[xshift=-1cm, dotted] (3.67,-1.5) node[circle,fill,inner sep=1pt](e){} -- (4.46,-1.5) 
node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{5}$](f){};

\draw[xshift=-1cm, dotted] (3.89,-1.9) node[circle,fill,inner sep=1pt, label=left:$-\frac{4}{7}$](g){} -- (4.46,-1.9) 
node[circle,fill,inner sep=1pt](h){};

\draw[xshift=-1cm, dotted] (3.89,-2.3) node[circle,fill,inner sep=1pt](i){} -- (4.33,-2.3) 
node[circle,fill,inner sep=1pt, label=right:$+\frac{4}{9}$](j){};

\draw[->, to path={-| (\tikztotarget)}, dotted]
  (b) edge (d) (c) edge (e) (f) edge (h) (g) edge (i);
     \end{tikzpicture}

\label{fig:numberline}
\caption[]{A number line demonstrating the convergence of Madhava's first method for the 
calculation of $\pi$. Adapted from (V. M. Jamkar) \cite{jamkar_2018}}
\end{figure}
